### 2022-04-05 tables/rep_table_1.tex
cleaning, added Table 1 and Table 2

\begin{table}[H]

\caption{\label{tab:tab:replicatetable2}Replication of Table 1}
\centering
\begin{tabular}[t]{lrr}
\toprule
Variable & Standard Deviation of State Means & Mean of Within-State Standard Deviations\\
\midrule
\cellcolor{gray!6}{Violent Crime Rate} & \cellcolor{gray!6}{306.62} & \cellcolor{gray!6}{78.19}\\
Property Crime Rate & 1135.99 & 422.15\\
\cellcolor{gray!6}{Murder Crime Rate} & \cellcolor{gray!6}{6.19} & \cellcolor{gray!6}{1.50}\\
Rape Crime Rate & 13.26 & 6.37\\
\cellcolor{gray!6}{Robbery Crime Rate} & \cellcolor{gray!6}{171.87} & \cellcolor{gray!6}{28.81}\\
\addlinespace
Assault Crime Rate & 144.28 & 57.62\\
\cellcolor{gray!6}{Burglary Crime Rate} & \cellcolor{gray!6}{372.52} & \cellcolor{gray!6}{181.39}\\
Larceny Crime Rate & 705.73 & 260.55\\
\cellcolor{gray!6}{Autotheft Crime Rate} & \cellcolor{gray!6}{210.57} & \cellcolor{gray!6}{82.24}\\
Violent Arrest Rate & 10.38 & 10.54\\
\addlinespace
\cellcolor{gray!6}{Property Arrest Rate} & \cellcolor{gray!6}{3.07} & \cellcolor{gray!6}{3.23}\\
Murder Arrest Rate & 17.26 & 29.23\\
\cellcolor{gray!6}{Rape Arrest Rate} & \cellcolor{gray!6}{11.15} & \cellcolor{gray!6}{10.65}\\
Assault Arrest Rate & 11.42 & 10.81\\
\cellcolor{gray!6}{Robery Arrest Rate} & \cellcolor{gray!6}{8.83} & \cellcolor{gray!6}{8.09}\\
\addlinespace
Burglary Arrest Rate & 3.08 & 3.20\\
\cellcolor{gray!6}{Larceny Arrest Rate} & \cellcolor{gray!6}{3.42} & \cellcolor{gray!6}{3.49}\\
Autotheft Arrest Rate & 10.67 & 31.62\\
\bottomrule
\end{tabular}
\end{table}