### 2022-04-05 tables/rep_table_2.tex

\begin{table}[H]

\caption{\label{tab:tab:replicatetable2}Replication of Table 2}
\centering
\begin{tabular}[t]{lrrr}
\toprule
Variable & N & Mean & Standard Deviation\\
\midrule
\cellcolor{gray!6}{Shalll} & \cellcolor{gray!6}{816} & \cellcolor{gray!6}{0.19} & \cellcolor{gray!6}{0.39}\\
Violent Arrest Rate & 802 & 41.09 & 22.20\\
\cellcolor{gray!6}{Property Arrest Rate} & \cellcolor{gray!6}{809} & \cellcolor{gray!6}{16.92} & \cellcolor{gray!6}{4.68}\\
Murder Arrest Rate & 806 & 91.30 & 55.94\\
\cellcolor{gray!6}{Rape Arrest Rate} & \cellcolor{gray!6}{799} & \cellcolor{gray!6}{41.02} & \cellcolor{gray!6}{17.39}\\
\addlinespace
Assault Arrest Rate & 809 & 44.62 & 16.98\\
\cellcolor{gray!6}{Robery Arrest Rate} & \cellcolor{gray!6}{808} & \cellcolor{gray!6}{31.46} & \cellcolor{gray!6}{13.59}\\
Burglary Arrest Rate & 809 & 13.80 & 4.57\\
\cellcolor{gray!6}{Larceny Arrest Rate} & \cellcolor{gray!6}{809} & \cellcolor{gray!6}{18.54} & \cellcolor{gray!6}{5.20}\\
Autotheft Arrest Rate & 808 & 22.35 & 37.61\\
\addlinespace
\cellcolor{gray!6}{Violent Crime Rate} & \cellcolor{gray!6}{816} & \cellcolor{gray!6}{483.93} & \cellcolor{gray!6}{318.94}\\
Property Crime Rate & 816 & 4618.34 & 1210.46\\
\cellcolor{gray!6}{Murder Crime Rate} & \cellcolor{gray!6}{816} & \cellcolor{gray!6}{7.77} & \cellcolor{gray!6}{6.88}\\
Rape Crime Rate & 816 & 33.98 & 15.07\\
\cellcolor{gray!6}{Robbery Crime Rate} & \cellcolor{gray!6}{816} & \cellcolor{gray!6}{163.42} & \cellcolor{gray!6}{176.25}\\
\addlinespace
Assault Crime Rate & 816 & 278.76 & 159.65\\
\cellcolor{gray!6}{Burglary Crime Rate} & \cellcolor{gray!6}{816} & \cellcolor{gray!6}{1239.34} & \cellcolor{gray!6}{417.76}\\
Larceny Crime Rate & 816 & 2968.71 & 751.02\\
\cellcolor{gray!6}{Autotheft Crime Rate} & \cellcolor{gray!6}{816} & \cellcolor{gray!6}{410.30} & \cellcolor{gray!6}{231.15}\\
Personal Income Rpc & 816 & 9351.82 & 4689.70\\
\addlinespace
\cellcolor{gray!6}{Unemployment Insurance Rpc} & \cellcolor{gray!6}{816} & \cellcolor{gray!6}{50.02} & \cellcolor{gray!6}{38.08}\\
Income Maintenance Rpc & 816 & 115.28 & 70.95\\
\cellcolor{gray!6}{Retirement Payments Rpc} & \cellcolor{gray!6}{816} & \cellcolor{gray!6}{1002.23} & \cellcolor{gray!6}{546.47}\\
State Population & 816 & 4646787.34 & 5010349.87\\
\bottomrule
\end{tabular}
\end{table}