### 2022-04-06 tables/rep_table_2a.tex
update draft of twfe analysis, draft export table using etable.

\begin{table}[H]

\caption{\label{tab:tab:replicatetable2a}Replication of Table 2 (Panel A)}
\centering
\begin{tabular}[t]{lrrr}
\toprule
Variable & N & Mean & Standard Deviation\\
\midrule
\cellcolor{gray!6}{Shalll} & \cellcolor{gray!6}{816} & \cellcolor{gray!6}{0.1912} & \cellcolor{gray!6}{0.3935}\\
Violent Arrest Rate & 802 & 41.0906 & 22.2036\\
\cellcolor{gray!6}{Property Arrest Rate} & \cellcolor{gray!6}{809} & \cellcolor{gray!6}{16.9180} & \cellcolor{gray!6}{4.6770}\\
Murder Arrest Rate & 806 & 91.2989 & 55.9428\\
\cellcolor{gray!6}{Rape Arrest Rate} & \cellcolor{gray!6}{799} & \cellcolor{gray!6}{41.0231} & \cellcolor{gray!6}{17.3887}\\
\addlinespace
Assault Arrest Rate & 809 & 44.6250 & 16.9783\\
\cellcolor{gray!6}{Robery Arrest Rate} & \cellcolor{gray!6}{808} & \cellcolor{gray!6}{31.4581} & \cellcolor{gray!6}{13.5928}\\
Burglary Arrest Rate & 809 & 13.8044 & 4.5712\\
\cellcolor{gray!6}{Larceny Arrest Rate} & \cellcolor{gray!6}{809} & \cellcolor{gray!6}{18.5372} & \cellcolor{gray!6}{5.1961}\\
Autotheft Arrest Rate & 808 & 22.3455 & 37.6114\\
\addlinespace
\cellcolor{gray!6}{Violent Crime Rate} & \cellcolor{gray!6}{816} & \cellcolor{gray!6}{483.9260} & \cellcolor{gray!6}{318.9425}\\
Property Crime Rate & 816 & 4618.3394 & 1210.4646\\
\cellcolor{gray!6}{Murder Crime Rate} & \cellcolor{gray!6}{816} & \cellcolor{gray!6}{7.7683} & \cellcolor{gray!6}{6.8817}\\
Rape Crime Rate & 816 & 33.9818 & 15.0721\\
\cellcolor{gray!6}{Robbery Crime Rate} & \cellcolor{gray!6}{816} & \cellcolor{gray!6}{163.4208} & \cellcolor{gray!6}{176.2506}\\
\addlinespace
Assault Crime Rate & 816 & 278.7551 & 159.6495\\
\cellcolor{gray!6}{Burglary Crime Rate} & \cellcolor{gray!6}{816} & \cellcolor{gray!6}{1239.3364} & \cellcolor{gray!6}{417.7576}\\
Larceny Crime Rate & 816 & 2968.7079 & 751.0234\\
\cellcolor{gray!6}{Autotheft Crime Rate} & \cellcolor{gray!6}{816} & \cellcolor{gray!6}{410.2951} & \cellcolor{gray!6}{231.1537}\\
Personal Income Rpc & 816 & 9351.8205 & 4689.7012\\
\addlinespace
\cellcolor{gray!6}{Unemployment Insurance Rpc} & \cellcolor{gray!6}{816} & \cellcolor{gray!6}{50.0187} & \cellcolor{gray!6}{38.0808}\\
Income Maintenance Rpc & 816 & 115.2756 & 70.9528\\
\cellcolor{gray!6}{Retirement Payments Rpc} & \cellcolor{gray!6}{816} & \cellcolor{gray!6}{1002.2257} & \cellcolor{gray!6}{546.4679}\\
State Population & 816 & 4646787.3419 & 5010349.8734\\
\cellcolor{gray!6}{Density} & \cellcolor{gray!6}{816} & \cellcolor{gray!6}{355.9729} & \cellcolor{gray!6}{1408.2501}\\
\bottomrule
\end{tabular}
\end{table}