### 2022-04-06 tables/rep_table_2b.tex

\begin{table}[H]

\caption{\label{tab:tab:replicatetable2b}Replication of Table 2 (Panel B)}
\centering
\begin{tabular}[t]{lrrr}
\toprule
Variable & N & Mean & Standard Deviation\\
\midrule
\cellcolor{gray!6}{White Male 1019} & \cellcolor{gray!6}{816} & \cellcolor{gray!6}{0.0672} & \cellcolor{gray!6}{0.0151}\\
Black Male 1019 & 816 & 0.0102 & 0.0112\\
\cellcolor{gray!6}{Other Male 1019} & \cellcolor{gray!6}{816} & \cellcolor{gray!6}{0.0036} & \cellcolor{gray!6}{0.0076}\\
White Female 1019 & 816 & 0.0641 & 0.0149\\
\cellcolor{gray!6}{Black Female 1019} & \cellcolor{gray!6}{816} & \cellcolor{gray!6}{0.0101} & \cellcolor{gray!6}{0.0114}\\
\addlinespace
Other Female 1019 & 816 & 0.0035 & 0.0073\\
\cellcolor{gray!6}{White Male 2029} & \cellcolor{gray!6}{816} & \cellcolor{gray!6}{0.0741} & \cellcolor{gray!6}{0.0120}\\
Black Male 2029 & 816 & 0.0096 & 0.0104\\
\cellcolor{gray!6}{Other Male 2029} & \cellcolor{gray!6}{816} & \cellcolor{gray!6}{0.0035} & \cellcolor{gray!6}{0.0072}\\
White Female 2029 & 816 & 0.0726 & 0.0123\\
\addlinespace
\cellcolor{gray!6}{Black Female 2029} & \cellcolor{gray!6}{816} & \cellcolor{gray!6}{0.0104} & \cellcolor{gray!6}{0.0123}\\
Other Female 2029 & 816 & 0.0036 & 0.0073\\
\cellcolor{gray!6}{White Male 3039} & \cellcolor{gray!6}{816} & \cellcolor{gray!6}{0.0663} & \cellcolor{gray!6}{0.0118}\\
Black Male 3039 & 816 & 0.0071 & 0.0082\\
\cellcolor{gray!6}{Other Male 3039} & \cellcolor{gray!6}{816} & \cellcolor{gray!6}{0.0030} & \cellcolor{gray!6}{0.0066}\\
\addlinespace
White Female 3039 & 816 & 0.0658 & 0.0117\\
\cellcolor{gray!6}{Black Female 3039} & \cellcolor{gray!6}{816} & \cellcolor{gray!6}{0.0081} & \cellcolor{gray!6}{0.0098}\\
Other Female 3039 & 816 & 0.0033 & 0.0070\\
\cellcolor{gray!6}{White Male 4049} & \cellcolor{gray!6}{816} & \cellcolor{gray!6}{0.0480} & \cellcolor{gray!6}{0.0091}\\
Black Male 4049 & 816 & 0.0046 & 0.0056\\
\addlinespace
\cellcolor{gray!6}{Other Male 4049} & \cellcolor{gray!6}{816} & \cellcolor{gray!6}{0.0020} & \cellcolor{gray!6}{0.0047}\\
White Female 4049 & 816 & 0.0483 & 0.0090\\
\cellcolor{gray!6}{Black Female 4049} & \cellcolor{gray!6}{816} & \cellcolor{gray!6}{0.0055} & \cellcolor{gray!6}{0.0068}\\
Other Female 4049 & 816 & 0.0022 & 0.0051\\
\cellcolor{gray!6}{White Male 5064} & \cellcolor{gray!6}{816} & \cellcolor{gray!6}{0.0580} & \cellcolor{gray!6}{0.0105}\\
\addlinespace
Black Male 5064 & 816 & 0.0052 & 0.0069\\
\cellcolor{gray!6}{Other Male 5064} & \cellcolor{gray!6}{816} & \cellcolor{gray!6}{0.0019} & \cellcolor{gray!6}{0.0061}\\
White Female 5064 & 816 & 0.0625 & 0.0123\\
\cellcolor{gray!6}{Black Female 5064} & \cellcolor{gray!6}{816} & \cellcolor{gray!6}{0.0065} & \cellcolor{gray!6}{0.0089}\\
Other Female 5064 & 816 & 0.0023 & 0.0069\\
\addlinespace
\cellcolor{gray!6}{White Male 65o} & \cellcolor{gray!6}{816} & \cellcolor{gray!6}{0.0426} & \cellcolor{gray!6}{0.0114}\\
Black Male 65o & 816 & 0.0035 & 0.0048\\
\cellcolor{gray!6}{Other Male 65o} & \cellcolor{gray!6}{816} & \cellcolor{gray!6}{0.0012} & \cellcolor{gray!6}{0.0046}\\
White Female 65o & 816 & 0.0623 & 0.0167\\
\cellcolor{gray!6}{Black Female 65o} & \cellcolor{gray!6}{816} & \cellcolor{gray!6}{0.0054} & \cellcolor{gray!6}{0.0078}\\
\addlinespace
Other Female 65o & 816 & 0.0014 & 0.0048\\
\bottomrule
\end{tabular}
\end{table}