### 2022-04-06 tables/rep_table_3b.tex

\begin{table}[htbp]
   \caption{\label{tab:replicatetable3b} Replication of Table 3 Panel B : Fixed Effects Regressions}
   \centering
   \small
   \begin{tabular}{lccc}
      \tabularnewline \midrule \midrule
      Dependent Variables:     & rape\_crime\_rate\_log    & assault\_crime\_rate\_log    & robbery\_crime\_rate\_log\\     
      Model:                   & (1)                       & (2)                          & (3)\\  
      \midrule
      \emph{Variables}\\
      shalll                   & -0.0340                   & -0.1004$^{**}$               & -0.0532\\   
                               & (0.0395)                  & (0.0424)                     & (0.0439)\\   
      Relevant\_Arrest\_Rate   & -0.0006                   & -0.0028$^{***}$              & -0.0014\\   
                               & (0.0005)                  & (0.0009)                     & (0.0009)\\   
      \midrule
      \emph{Fixed-effects}\\
      state                    & Yes                       & Yes                          & Yes\\  
      year                     & Yes                       & Yes                          & Yes\\  
      \midrule
      \emph{Fit statistics}\\
      Observations             & 799                       & 809                          & 808\\  
      R$^2$                    & 0.94240                   & 0.96632                      & 0.98389\\  
      Within R$^2$             & 0.52351                   & 0.51645                      & 0.50471\\  
      \midrule \midrule
      \multicolumn{4}{l}{\emph{Clustered (state) standard-errors in parentheses}}\\
      \multicolumn{4}{l}{\emph{Signif. Codes: ***: 0.01, **: 0.05, *: 0.1}}\\
   \end{tabular}

   \par \raggedright 
   Control variables ommited from table, 
                           though they were included in the analysis. 
                           Consistent with the original paper, 
                           control variables include: density, personal\_income\_rpc, unemployment\_insurance\_rpc, income\_maintenance\_rpc, retirement\_payments\_rpc, state\_population, ppwm1019, ppbm1019, ppnm1019, ppwf1019, ppbf1019, ppnf1019, ppwm2029, ppbm2029, ppnm2029, ppwf2029, ppbf2029, ppnf2029, ppwm3039, ppbm3039, ppnm3039, ppwf3039, ppbf3039, ppnf3039, ppwm4049, ppbm4049, ppnm4049, ppwf4049, ppbf4049, ppnf4049, ppwm5064, ppbm5064, ppnm5064, ppwf5064, ppbf5064, ppnf5064, ppwm65o, ppbm65o, ppnm65o, ppwf65o, ppbf65o, ppnf65o.
\end{table}