### 2022-04-07 tables/bacon_decomposition.tex
bacon decomp

\begin{table}[H]

\caption{\label{tab:tab:bacondecomposition}Bacon Decomposition}
\centering
\begin{tabular}[t]{lrrrrrrrrr}
\toprule
Type & Violent crime rate log & Property crime rate log & Murder crime rate log & Rape crime rate log & Assault crime rate log & Robbery crime rate log & Burglary crime rate log & Larceny crime rate log & Autotheft crime rate log\\
\midrule
\cellcolor{gray!6}{Earlier vs Later Treated} & \cellcolor{gray!6}{0.0051705} & \cellcolor{gray!6}{-0.0007191} & \cellcolor{gray!6}{0.0054525} & \cellcolor{gray!6}{-0.0026425} & \cellcolor{gray!6}{0.0079628} & \cellcolor{gray!6}{0.0073682} & \cellcolor{gray!6}{-0.0023226} & \cellcolor{gray!6}{-0.0004160} & \cellcolor{gray!6}{0.0056843}\\
Later vs Always Treated & -0.0044690 & 0.0085564 & -0.0012542 & -0.0305273 & 0.0009728 & 0.0174822 & 0.0048413 & 0.0080003 & 0.0336117\\
\cellcolor{gray!6}{Later vs Earlier Treated} & \cellcolor{gray!6}{-0.0017883} & \cellcolor{gray!6}{0.0001507} & \cellcolor{gray!6}{0.0000418} & \cellcolor{gray!6}{-0.0019283} & \cellcolor{gray!6}{-0.0034427} & \cellcolor{gray!6}{0.0020947} & \cellcolor{gray!6}{-0.0013013} & \cellcolor{gray!6}{0.0004859} & \cellcolor{gray!6}{0.0020305}\\
Treated vs Untreated & -0.0839315 & 0.0210569 & -0.0415959 & 0.0030423 & -0.1375213 & -0.0100514 & 0.0064291 & 0.0280699 & 0.0262756\\
\cellcolor{gray!6}{Total TWFE} & \cellcolor{gray!6}{-0.0850183} & \cellcolor{gray!6}{0.0290449} & \cellcolor{gray!6}{-0.0373558} & \cellcolor{gray!6}{-0.0320558} & \cellcolor{gray!6}{-0.1320284} & \cellcolor{gray!6}{0.0168936} & \cellcolor{gray!6}{0.0076465} & \cellcolor{gray!6}{0.0361400} & \cellcolor{gray!6}{0.0676020}\\
\bottomrule
\end{tabular}
\end{table}