### 2022-04-07 tables/bacon_decompositiona.tex

\begin{table}[H]

\caption{\label{tab:tab:bacondecompositiona}Bacon Decomposition: Panel A}
\centering
\begin{tabular}[t]{lrrr}
\toprule
Type & Violent crime rate log & Property crime rate log & Murder crime rate log\\
\midrule
\cellcolor{gray!6}{Earlier vs Later Treated} & \cellcolor{gray!6}{0.0051705} & \cellcolor{gray!6}{-0.0007191} & \cellcolor{gray!6}{0.0054525}\\
Later vs Always Treated & -0.0044690 & 0.0085564 & -0.0012542\\
\cellcolor{gray!6}{Later vs Earlier Treated} & \cellcolor{gray!6}{-0.0017883} & \cellcolor{gray!6}{0.0001507} & \cellcolor{gray!6}{0.0000418}\\
Treated vs Untreated & -0.0839315 & 0.0210569 & -0.0415959\\
\cellcolor{gray!6}{Total TWFE} & \cellcolor{gray!6}{-0.0850183} & \cellcolor{gray!6}{0.0290449} & \cellcolor{gray!6}{-0.0373558}\\
\bottomrule
\end{tabular}
\end{table}