### 2022-04-07 tables/bacon_decompositionb.tex

\begin{table}[H]

\caption{\label{tab:tab:bacondecompositionb}Bacon Decomposition: Panel B}
\centering
\begin{tabular}[t]{lrrr}
\toprule
Type & Rape crime rate log & Assault crime rate log & Robbery crime rate log\\
\midrule
\cellcolor{gray!6}{Earlier vs Later Treated} & \cellcolor{gray!6}{-0.0026425} & \cellcolor{gray!6}{0.0079628} & \cellcolor{gray!6}{0.0073682}\\
Later vs Always Treated & -0.0305273 & 0.0009728 & 0.0174822\\
\cellcolor{gray!6}{Later vs Earlier Treated} & \cellcolor{gray!6}{-0.0019283} & \cellcolor{gray!6}{-0.0034427} & \cellcolor{gray!6}{0.0020947}\\
Treated vs Untreated & 0.0030423 & -0.1375213 & -0.0100514\\
\cellcolor{gray!6}{Total TWFE} & \cellcolor{gray!6}{-0.0320558} & \cellcolor{gray!6}{-0.1320284} & \cellcolor{gray!6}{0.0168936}\\
\bottomrule
\end{tabular}
\end{table}