### 2022-04-07 tables/bacon_decompositionc.tex

\begin{table}[H]

\caption{\label{tab:tab:bacondecompositionc}Bacon Decomposition: Panel C}
\centering
\begin{tabular}[t]{lrrr}
\toprule
Type & Burglary crime rate log & Larceny crime rate log & Autotheft crime rate log\\
\midrule
\cellcolor{gray!6}{Earlier vs Later Treated} & \cellcolor{gray!6}{-0.0023226} & \cellcolor{gray!6}{-0.0004160} & \cellcolor{gray!6}{0.0056843}\\
Later vs Always Treated & 0.0048413 & 0.0080003 & 0.0336117\\
\cellcolor{gray!6}{Later vs Earlier Treated} & \cellcolor{gray!6}{-0.0013013} & \cellcolor{gray!6}{0.0004859} & \cellcolor{gray!6}{0.0020305}\\
Treated vs Untreated & 0.0064291 & 0.0280699 & 0.0262756\\
\cellcolor{gray!6}{Total TWFE} & \cellcolor{gray!6}{0.0076465} & \cellcolor{gray!6}{0.0361400} & \cellcolor{gray!6}{0.0676020}\\
\bottomrule
\end{tabular}
\end{table}