### 2022-04-20 tables/bacon_decomposition_Assault.tex
fix bacon decomp

\begin{table}[H]

\caption{\label{tab:tab:bacondecompositionAssault}Bacon Decomposition: Assault Crime Rate Log}
\centering
\begin{tabular}[t]{lrrr}
\toprule
Type & Average Estimate & Group Weight & Weighted Estimate\\
\midrule
\cellcolor{gray!6}{Earlier vs Later Treated} & \cellcolor{gray!6}{0.1487745} & \cellcolor{gray!6}{0.0683810} & \cellcolor{gray!6}{0.0079628}\\
Later vs Always Treated & -0.0314314 & 0.1589397 & 0.0009728\\
\cellcolor{gray!6}{Later vs Earlier Treated} & \cellcolor{gray!6}{-0.0174274} & \cellcolor{gray!6}{0.0233921} & \cellcolor{gray!6}{-0.0034427}\\
Treated vs Untreated & -0.2207888 & 0.7492871 & -0.1375213\\
\cellcolor{gray!6}{Total TWFE} & \cellcolor{gray!6}{NaN} & \cellcolor{gray!6}{NaN} & \cellcolor{gray!6}{-0.1320284}\\
\bottomrule
\end{tabular}
\end{table}