### 2022-04-20 tables/bacon_decomposition_Autotheft.tex

\begin{table}[H]

\caption{\label{tab:tab:bacondecompositionAutotheft}Bacon Decomposition: Autotheft Crime Rate Log}
\centering
\begin{tabular}[t]{lrrr}
\toprule
Type & Average Estimate & Group Weight & Weighted Estimate\\
\midrule
\cellcolor{gray!6}{Earlier vs Later Treated} & \cellcolor{gray!6}{0.0887179} & \cellcolor{gray!6}{0.0683810} & \cellcolor{gray!6}{0.0056843}\\
Later vs Always Treated & 0.1885142 & 0.1589397 & 0.0336117\\
\cellcolor{gray!6}{Later vs Earlier Treated} & \cellcolor{gray!6}{0.0994063} & \cellcolor{gray!6}{0.0233921} & \cellcolor{gray!6}{0.0020305}\\
Treated vs Untreated & 0.0086094 & 0.7492871 & 0.0262756\\
\cellcolor{gray!6}{Total TWFE} & \cellcolor{gray!6}{NaN} & \cellcolor{gray!6}{NaN} & \cellcolor{gray!6}{0.0676020}\\
\bottomrule
\end{tabular}
\end{table}