### 2022-04-20 tables/bacon_decomposition_Burglary.tex

\begin{table}[H]

\caption{\label{tab:tab:bacondecompositionBurglary}Bacon Decomposition: Burglary Crime Rate Log}
\centering
\begin{tabular}[t]{lrrr}
\toprule
Type & Average Estimate & Group Weight & Weighted Estimate\\
\midrule
\cellcolor{gray!6}{Earlier vs Later Treated} & \cellcolor{gray!6}{-0.0146061} & \cellcolor{gray!6}{0.0683810} & \cellcolor{gray!6}{-0.0023226}\\
Later vs Always Treated & 0.0169005 & 0.1589397 & 0.0048413\\
\cellcolor{gray!6}{Later vs Earlier Treated} & \cellcolor{gray!6}{-0.0157802} & \cellcolor{gray!6}{0.0233921} & \cellcolor{gray!6}{-0.0013013}\\
Treated vs Untreated & -0.0068820 & 0.7492871 & 0.0064291\\
\cellcolor{gray!6}{Total TWFE} & \cellcolor{gray!6}{NaN} & \cellcolor{gray!6}{NaN} & \cellcolor{gray!6}{0.0076465}\\
\bottomrule
\end{tabular}
\end{table}