### 2022-04-20 tables/bacon_decomposition_Larceny.tex

\begin{table}[H]

\caption{\label{tab:tab:bacondecompositionLarceny}Bacon Decomposition: Larceny Crime Rate Log}
\centering
\begin{tabular}[t]{lrrr}
\toprule
Type & Average Estimate & Group Weight & Weighted Estimate\\
\midrule
\cellcolor{gray!6}{Earlier vs Later Treated} & \cellcolor{gray!6}{0.0088734} & \cellcolor{gray!6}{0.0683810} & \cellcolor{gray!6}{-0.0004160}\\
Later vs Always Treated & 0.0477985 & 0.1589397 & 0.0080003\\
\cellcolor{gray!6}{Later vs Earlier Treated} & \cellcolor{gray!6}{0.0616798} & \cellcolor{gray!6}{0.0233921} & \cellcolor{gray!6}{0.0004859}\\
Treated vs Untreated & 0.0346292 & 0.7492871 & 0.0280699\\
\cellcolor{gray!6}{Total TWFE} & \cellcolor{gray!6}{NaN} & \cellcolor{gray!6}{NaN} & \cellcolor{gray!6}{0.0361400}\\
\bottomrule
\end{tabular}
\end{table}