### 2022-04-20 tables/bacon_decomposition_Murder.tex

\begin{table}[H]

\caption{\label{tab:tab:bacondecompositionMurder}Bacon Decomposition: Murder Crime Rate Log}
\centering
\begin{tabular}[t]{lrrr}
\toprule
Type & Average Estimate & Group Weight & Weighted Estimate\\
\midrule
\cellcolor{gray!6}{Earlier vs Later Treated} & \cellcolor{gray!6}{0.0545024} & \cellcolor{gray!6}{0.0683810} & \cellcolor{gray!6}{0.0054525}\\
Later vs Always Treated & -0.0384771 & 0.1589397 & -0.0012542\\
\cellcolor{gray!6}{Later vs Earlier Treated} & \cellcolor{gray!6}{0.0188246} & \cellcolor{gray!6}{0.0233921} & \cellcolor{gray!6}{0.0000418}\\
Treated vs Untreated & -0.0848318 & 0.7492871 & -0.0415959\\
\cellcolor{gray!6}{Total TWFE} & \cellcolor{gray!6}{NaN} & \cellcolor{gray!6}{NaN} & \cellcolor{gray!6}{-0.0373558}\\
\bottomrule
\end{tabular}
\end{table}