### 2022-04-20 tables/bacon_decomposition_Property.tex

\begin{table}[H]

\caption{\label{tab:tab:bacondecompositionProperty}Bacon Decomposition: Property Crime Rate Log}
\centering
\begin{tabular}[t]{lrrr}
\toprule
Type & Average Estimate & Group Weight & Weighted Estimate\\
\midrule
\cellcolor{gray!6}{Earlier vs Later Treated} & \cellcolor{gray!6}{0.0036098} & \cellcolor{gray!6}{0.0683810} & \cellcolor{gray!6}{-0.0007191}\\
Later vs Always Treated & 0.0504242 & 0.1589397 & 0.0085564\\
\cellcolor{gray!6}{Later vs Earlier Treated} & \cellcolor{gray!6}{0.0490119} & \cellcolor{gray!6}{0.0233921} & \cellcolor{gray!6}{0.0001507}\\
Treated vs Untreated & 0.0238073 & 0.7492871 & 0.0210569\\
\cellcolor{gray!6}{Total TWFE} & \cellcolor{gray!6}{NaN} & \cellcolor{gray!6}{NaN} & \cellcolor{gray!6}{0.0290449}\\
\bottomrule
\end{tabular}
\end{table}