### 2022-04-20 tables/bacon_decomposition_Rape.tex

\begin{table}[H]

\caption{\label{tab:tab:bacondecompositionRape}Bacon Decomposition: Rape Crime Rate Log}
\centering
\begin{tabular}[t]{lrrr}
\toprule
Type & Average Estimate & Group Weight & Weighted Estimate\\
\midrule
\cellcolor{gray!6}{Earlier vs Later Treated} & \cellcolor{gray!6}{-0.0264193} & \cellcolor{gray!6}{0.0683810} & \cellcolor{gray!6}{-0.0026425}\\
Later vs Always Treated & -0.2001560 & 0.1589397 & -0.0305273\\
\cellcolor{gray!6}{Later vs Earlier Treated} & \cellcolor{gray!6}{-0.0106483} & \cellcolor{gray!6}{0.0233921} & \cellcolor{gray!6}{-0.0019283}\\
Treated vs Untreated & -0.0030557 & 0.7492871 & 0.0030423\\
\cellcolor{gray!6}{Total TWFE} & \cellcolor{gray!6}{NaN} & \cellcolor{gray!6}{NaN} & \cellcolor{gray!6}{-0.0320558}\\
\bottomrule
\end{tabular}
\end{table}