### 2022-04-20 tables/bacon_decomposition_Robbery.tex

\begin{table}[H]

\caption{\label{tab:tab:bacondecompositionRobbery}Bacon Decomposition: Robbery Crime Rate Log}
\centering
\begin{tabular}[t]{lrrr}
\toprule
Type & Average Estimate & Group Weight & Weighted Estimate\\
\midrule
\cellcolor{gray!6}{Earlier vs Later Treated} & \cellcolor{gray!6}{0.0961047} & \cellcolor{gray!6}{0.0683810} & \cellcolor{gray!6}{0.0073682}\\
Later vs Always Treated & 0.0940074 & 0.1589397 & 0.0174822\\
\cellcolor{gray!6}{Later vs Earlier Treated} & \cellcolor{gray!6}{0.1505417} & \cellcolor{gray!6}{0.0233921} & \cellcolor{gray!6}{0.0020947}\\
Treated vs Untreated & -0.0263060 & 0.7492871 & -0.0100514\\
\cellcolor{gray!6}{Total TWFE} & \cellcolor{gray!6}{NaN} & \cellcolor{gray!6}{NaN} & \cellcolor{gray!6}{0.0168936}\\
\bottomrule
\end{tabular}
\end{table}