### 2022-04-20 tables/cs.tex
updates to cs, draft section

\begin{table}[H]

\caption{\label{tab:tab:cs}Callaway and Sant'anna Overall ATTs}
\centering
\begin{tabular}[t]{lrr}
\toprule
Outcome Variable & Overall ATT & Std. Error\\
\midrule
\cellcolor{gray!6}{Violent Crime Rate Log} & \cellcolor{gray!6}{-0.0099595} & \cellcolor{gray!6}{0.0259180}\\
Property Crime Rate Log & 0.0129076 & 0.0128726\\
\cellcolor{gray!6}{Murder Crime Rate Log} & \cellcolor{gray!6}{-0.0485681} & \cellcolor{gray!6}{0.0266759}\\
Rape Crime Rate Log & 0.0200170 & 0.0284581\\
\cellcolor{gray!6}{Assault Crime Rate Log} & \cellcolor{gray!6}{0.0048334} & \cellcolor{gray!6}{0.0444671}\\
\addlinespace
Robbery Crime Rate Log & 0.0387509 & 0.0344164\\
\cellcolor{gray!6}{Burglary Crime Rate Log} & \cellcolor{gray!6}{-0.0157512} & \cellcolor{gray!6}{0.0161027}\\
Larceny Crime Rate Log & 0.0302561 & 0.0165051\\
\cellcolor{gray!6}{Autotheft Crime Rate Log} & \cellcolor{gray!6}{0.0107402} & \cellcolor{gray!6}{0.0427668}\\
\bottomrule
\end{tabular}
\end{table}