### 2022-04-20 tables/twfe.tex

\begin{table}[H]

\caption{\label{tab:tab:twfe}Two Way Fixed Effects Shall Issue Coefficients}
\centering
\begin{tabular}[t]{lrr}
\toprule
Outcome Variable & Coefficient & Std. Error\\
\midrule
\cellcolor{gray!6}{Violent Crime Rate Log} & \cellcolor{gray!6}{-0.0977525} & \cellcolor{gray!6}{0.0207145}\\
Property Crime Rate Log & -0.0072047 & 0.0136816\\
\cellcolor{gray!6}{Murder Crime Rate Log} & \cellcolor{gray!6}{-0.0506590} & \cellcolor{gray!6}{0.0385770}\\
Rape Crime Rate Log & -0.0339755 & 0.0271763\\
\cellcolor{gray!6}{Assault Crime Rate Log} & \cellcolor{gray!6}{-0.1004015} & \cellcolor{gray!6}{0.0270163}\\
\addlinespace
Robbery Crime Rate Log & -0.0532480 & 0.0311260\\
\cellcolor{gray!6}{Burglary Crime Rate Log} & \cellcolor{gray!6}{-0.0460697} & \cellcolor{gray!6}{0.0190139}\\
Larceny Crime Rate Log & 0.0033107 & 0.0138442\\
\cellcolor{gray!6}{Autotheft Crime Rate Log} & \cellcolor{gray!6}{-0.0089805} & \cellcolor{gray!6}{0.0282748}\\
\bottomrule
\end{tabular}
\end{table}