### 2022-04-26 tables/bacon_decomposition_Assault.tex
final push

\begin{table}[H]

\caption{\label{tab:tab:bacondecompositionAssault}Bacon Decomposition - Assault Crime Rate Log}
\centering
\begin{tabular}[t]{lrrr}
\toprule
Type & Average Estimate & Group Weight & Weighted Estimate\\
\midrule
\cellcolor{gray!6}{Earlier vs Later Treated} & \cellcolor{gray!6}{0.149} & \cellcolor{gray!6}{0.068} & \cellcolor{gray!6}{0.008}\\
Later vs Always Treated & -0.031 & 0.159 & 0.001\\
\cellcolor{gray!6}{Later vs Earlier Treated} & \cellcolor{gray!6}{-0.017} & \cellcolor{gray!6}{0.023} & \cellcolor{gray!6}{-0.003}\\
Treated vs Untreated & -0.221 & 0.749 & -0.138\\
\cellcolor{gray!6}{Total TWFE} & \cellcolor{gray!6}{NaN} & \cellcolor{gray!6}{NaN} & \cellcolor{gray!6}{-0.132}\\
\bottomrule
\end{tabular}
\end{table}