### 2022-04-26 tables/bacon_decomposition_Larceny.tex

\begin{table}[H]

\caption{\label{tab:tab:bacondecompositionLarceny}Bacon Decomposition - Larceny Crime Rate Log}
\centering
\begin{tabular}[t]{lrrr}
\toprule
Type & Average Estimate & Group Weight & Weighted Estimate\\
\midrule
\cellcolor{gray!6}{Earlier vs Later Treated} & \cellcolor{gray!6}{0.009} & \cellcolor{gray!6}{0.068} & \cellcolor{gray!6}{0.000}\\
Later vs Always Treated & 0.048 & 0.159 & 0.008\\
\cellcolor{gray!6}{Later vs Earlier Treated} & \cellcolor{gray!6}{0.062} & \cellcolor{gray!6}{0.023} & \cellcolor{gray!6}{0.000}\\
Treated vs Untreated & 0.035 & 0.749 & 0.028\\
\cellcolor{gray!6}{Total TWFE} & \cellcolor{gray!6}{NaN} & \cellcolor{gray!6}{NaN} & \cellcolor{gray!6}{0.036}\\
\bottomrule
\end{tabular}
\end{table}