### 2022-04-26 tables/bacon_decomposition_Property.tex

\begin{table}[H]

\caption{\label{tab:tab:bacondecompositionProperty}Bacon Decomposition - Property Crime Rate Log}
\centering
\begin{tabular}[t]{lrrr}
\toprule
Type & Average Estimate & Group Weight & Weighted Estimate\\
\midrule
\cellcolor{gray!6}{Earlier vs Later Treated} & \cellcolor{gray!6}{0.004} & \cellcolor{gray!6}{0.068} & \cellcolor{gray!6}{-0.001}\\
Later vs Always Treated & 0.050 & 0.159 & 0.009\\
\cellcolor{gray!6}{Later vs Earlier Treated} & \cellcolor{gray!6}{0.049} & \cellcolor{gray!6}{0.023} & \cellcolor{gray!6}{0.000}\\
Treated vs Untreated & 0.024 & 0.749 & 0.021\\
\cellcolor{gray!6}{Total TWFE} & \cellcolor{gray!6}{NaN} & \cellcolor{gray!6}{NaN} & \cellcolor{gray!6}{0.029}\\
\bottomrule
\end{tabular}
\end{table}