### 2022-04-26 tables/bacon_decomposition_Violent.tex

\begin{table}[H]

\caption{\label{tab:tab:bacondecompositionViolent}Bacon Decomposition - Violent Crime Rate Log}
\centering
\begin{tabular}[t]{lrrr}
\toprule
Type & Average Estimate & Group Weight & Weighted Estimate\\
\midrule
\cellcolor{gray!6}{Earlier vs Later Treated} & \cellcolor{gray!6}{0.100} & \cellcolor{gray!6}{0.068} & \cellcolor{gray!6}{0.005}\\
Later vs Always Treated & -0.060 & 0.159 & -0.004\\
\cellcolor{gray!6}{Later vs Earlier Treated} & \cellcolor{gray!6}{0.021} & \cellcolor{gray!6}{0.023} & \cellcolor{gray!6}{-0.002}\\
Treated vs Untreated & -0.142 & 0.749 & -0.084\\
\cellcolor{gray!6}{Total TWFE} & \cellcolor{gray!6}{NaN} & \cellcolor{gray!6}{NaN} & \cellcolor{gray!6}{-0.085}\\
\bottomrule
\end{tabular}
\end{table}