### 2022-04-26 tables/cs.tex

\begin{table}[H]

\caption{\label{tab:tab:cs}Callaway and Sant'anna Overall ATTs}
\centering
\begin{tabular}[t]{lrr}
\toprule
Outcome Variable & Overall ATT & Std. Error\\
\midrule
\cellcolor{gray!6}{Violent Crime Rate Log} & \cellcolor{gray!6}{-0.010} & \cellcolor{gray!6}{0.025}\\
Property Crime Rate Log & 0.013 & 0.012\\
\cellcolor{gray!6}{Murder Crime Rate Log} & \cellcolor{gray!6}{-0.049} & \cellcolor{gray!6}{0.027}\\
Rape Crime Rate Log & 0.020 & 0.028\\
\cellcolor{gray!6}{Assault Crime Rate Log} & \cellcolor{gray!6}{0.005} & \cellcolor{gray!6}{0.041}\\
\addlinespace
Robbery Crime Rate Log & 0.039 & 0.033\\
\cellcolor{gray!6}{Burglary Crime Rate Log} & \cellcolor{gray!6}{-0.016} & \cellcolor{gray!6}{0.016}\\
Larceny Crime Rate Log & 0.030 & 0.016\\
\cellcolor{gray!6}{Autotheft Crime Rate Log} & \cellcolor{gray!6}{0.011} & \cellcolor{gray!6}{0.041}\\
\bottomrule
\end{tabular}
\end{table}