### 2022-04-26 tables/twfe.tex

\begin{table}[H]

\caption{\label{tab:tab:twfe}Two Way Fixed Effects Shall Issue Coefficients}
\centering
\begin{tabular}[t]{lrr}
\toprule
Outcome Variable & Coefficient & Std. Error\\
\midrule
\cellcolor{gray!6}{Violent Crime Rate Log} & \cellcolor{gray!6}{-0.098} & \cellcolor{gray!6}{0.021}\\
Property Crime Rate Log & -0.007 & 0.014\\
\cellcolor{gray!6}{Murder Crime Rate Log} & \cellcolor{gray!6}{-0.051} & \cellcolor{gray!6}{0.039}\\
Rape Crime Rate Log & -0.034 & 0.027\\
\cellcolor{gray!6}{Assault Crime Rate Log} & \cellcolor{gray!6}{-0.100} & \cellcolor{gray!6}{0.027}\\
\addlinespace
Robbery Crime Rate Log & -0.053 & 0.031\\
\cellcolor{gray!6}{Burglary Crime Rate Log} & \cellcolor{gray!6}{-0.046} & \cellcolor{gray!6}{0.019}\\
Larceny Crime Rate Log & 0.003 & 0.014\\
\cellcolor{gray!6}{Autotheft Crime Rate Log} & \cellcolor{gray!6}{-0.009} & \cellcolor{gray!6}{0.028}\\
\bottomrule
\end{tabular}
\end{table}